---
title: 'Psych 5066: Homework #5'
author: YOUR NAME HERE
output: 
  html_document
---

# Instructions

Please complete this assignment using the RMarkdown file provided (<a href="homework-5.Rmd" download="homework-5.Rmd">download here</a>). Be sure to (1) include your name in the preamble, (2) save the file, including your last name **AND THE NAME OF YOUR AI** (e.g., "cooper_april_homework_5.Rmd"). When you turn in the assignment, include **_both_** the .Rmd and knitted .pdf files.

  - If you are in Psychology or Computational & Data Sciences: your AI is Jayde
  - If you are in any of the other programs: your AI is April

There will be empty gray sections below each question. Please put all R `CODE` within these gray chunks. They will behave as normal R code. If you are asked to explain something or write a few sentences, please do that *outside* of the gray chunks. That way, when you knit, your sentences will show up as normal text, but all of your code will be in a designated code chunk.

You will be able to retry questions to earn full credit, but here are some things to keep in mind:

1. You can only do this for questions that you attempted on the first try. If you skip a question, you will not be able to resubmit for full credit. 

2. For each question, the points are all-or-nothing, meaning you cannot receive partial credit if you correctly answer only some of the question (this includes the description/explanation of results).

3. Late assignments will receive 50% of the points earned, with no opportunity to resubmit your answers for a higher grade (unless otherwise discussed with Dr. Cooper, Jayde, or April). For example, if you correctly answer questions and earn 10 points but submit the assignment late, you will receive 5 points for the assignment. 

4. You are allowed to Google things. *However*, if you use code from an outside source (meaning not the lecture slides or the R help documentation), then you must provide a **working link** to the website you used. You don't need to use APA formatting or anything. Just a working url. You may discuss homework assignments with your classmates; however, it is important that you complete each assignment on your own and **do not simply copy someone else’s code**. If we believe one student has copied another’s work, both students will receive a 0 on the homework assignment and will not be allowed to resubmit the assignment for points.

5. In the instructions, when it says **_by hand_**, what we mean is that you should not use a function that is built into R to complete the task. For example, if I ask you to find the mean *by hand*, you should **not** use the `mean()` function. However, you can (and should!) use your variables. So for instance, if I ask you to find the `mean()` by hand, you may do something like `sum(variableX)/length(variableX)`. This would be an appropriate *by hand* calculation. (You don't need to type `(3 + 5 + 5 + 17) / 4`...you can still use your variables). 

**Data:** The questions in this assignment use the dataset referred to as `homework-bdi`. [Link to data here.](homework-bdi.csv) For the problems below, we are interested in investigating the efficacy of an intervention to reduce depressive symptoms in a randomly sampled, community-based sample of adults. The participants completed the Beck Depression Inventory (BDI), and the data file includes the participants' scores on that dimension. Additionally, you will need [this dataset](homework-bdi-part2.csv) to complete question 3. 

Good luck!

## Setup (1 pt)
I suggest including a setup chunk to load your dataset and make sure it is in a working data.frame format. You might consider loading the `tidyverse` package, if you wish (but not required).

```{r}

```

## Question 1 (6 pts total)
For now, ignore the `condition` column of your data.frame and only pay attention to the scores. You recently read a literature review that suggested that your study population has an average BDI score of 14.

### 1a) Write out the restricted and full models for the analysis, in your own words (e.g., NOT equations)


### 1b) Conduct the appropriate analysis to compare these models.
You must show all of your work! It's hard to see where you went wrong if we can't view what you calculated.

```{r}

```

### 1c) Calculate an appropriate effect size for the analysis

```{r}

```

## Question 2 (6 pts total)

Participants were randomly assigned into either the intervention group or a waitlist control groupl scores reflect particpants' BDI scores following the term of the intervention. In this dataset, 0 indicates participants in the control group, 1 indicates particpants in the intervention group. 

### 2a) Write out the restricted and full models for the analysis, in your own words (e.g., NOT equations)


### 2b) Conduct the appropriate analysis to compare these models.
You must show all of your work! It's hard to see where you went wrong if we can't view what you calculated.

```{r}

```

### 2c) Calculate an appropriate effect size for the analysis

```{r}

```

## Question 3 (8 pts total)

### 3a) Next, you need to combine the current dataset with the additional one. The new dataset includes new data with Condition 2. Merge the 2 datasets together to create a single file to complete the remaining questions. 

```{r}

```

### 3b) Now assume that you have 3 groups (0 -- Control, 1 -- Intervention A, 2 -- Intervention B). Write out the restricted and full models for the oneway ANOVA of whether any group differences occur, in your own words (e.g., NOT equations)


### 3c) Conduct the appropriate analysis to compare these models

```{r}

```

### 3d) Calculate an effect size measure for the overall ANOVA test of differences across groups

```{r}

```

## Question 4 (4 pts)

Provide an interpretation of each of the 3 analyses you just ran (3 total). Discuss how things changed when going from analysis 1 to analysis 2 (what happened to your $t$-statistic and effect size?).  
